---
title: "Appendix"
author: "Chris Perfitt"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,  # hide code
  message = FALSE,  # hide package messages
  warning = FALSE   # hide warnings
)
```

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(broom)
library(plm)
library(purrr)
library(stargazer)
library(ggforce)
library(ggrepel)
library(scales)
library(strucchange)
library(lmtest)
library(sandwich)
library(ARDL)
library(zoo)
library(dplyr)
library(tibble)
library(knitr)
library(car)
library(kableExtra)
```

```{r}
pop_raw <- read_csv(
  "/Users/chris/Desktop/CADPOP.csv",
  skip = 10,
  show_col_types = FALSE
)

pop_raw <- pop_raw %>%
  mutate(across(everything(), as.character))

cad_pop <- pop_raw %>%
  filter(`Age group 3 6` == "All ages") %>%
  pivot_longer(
    cols      = -`Age group 3 6`,
    names_to  = "Year",
    values_to = "Population"
  ) %>%
  mutate(
    Year       = as.integer(Year),
    Population = readr::parse_number(Population),
    COUNTRY    = "Canada"
  ) %>%
  select(COUNTRY, Year, Population)
```

```{r}
mfp_raw <- read_csv(
  "/Users/chris/Desktop/CADMFP.csv",
  skip = 10,
  show_col_types = FALSE
)

mfp_raw <- mfp_raw %>%
  mutate(across(everything(), as.character))

cad_mfp <- mfp_raw %>%
  filter(`North American Industry Classification System (NAICS) 2` ==
           "Total economy 3") %>%
  pivot_longer(
    cols      = -`North American Industry Classification System (NAICS) 2`,
    names_to  = "quarter",
    values_to = "index"
  ) %>%
  mutate(
    index = readr::parse_number(index),
    Year  = as.integer(sub(".* ", "", quarter))
  ) %>%
  group_by(Year) %>%
  summarise(
    MFP_index = mean(index, na.rm = TRUE),
    .groups   = "drop"
  ) %>%
  mutate(COUNTRY = "Canada") %>%
  select(COUNTRY, Year, MFP_index)
```

```{r}
cad_gdp_raw <- read_csv(
  "/Users/chris/Desktop/CADGDP.csv",
  show_col_types = FALSE
)

cad_gdp <- cad_gdp_raw %>%
  # Restrict to the series you actually want
  filter(
    GEO      == "Canada",
    Estimates == "Gross domestic product at market prices",
    Prices   == "Chained (2017) dollars"
  ) %>%
  transmute(
    COUNTRY   = GEO,
    Year      = as.integer(REF_DATE),
    gdp_level = as.numeric(VALUE)
  )
```

```{r}
debt_raw <- read_csv(
  "/Users/chris/Desktop/CAD DEBT DATA.csv",
  col_types = cols(.default = "c"),
  show_col_types = FALSE
)

# All columns that are 4-digit years
year_cols <- grep("^[0-9]{4}$", names(debt_raw), value = TRUE)

# If there are multiple countries/indicators, filter here. If it's just Canada, this is harmless.
debt_ca_raw <- debt_raw %>%
  filter(COUNTRY == "Canada")

cad_debt <- debt_ca_raw %>%
  pivot_longer(
    cols      = all_of(year_cols),
    names_to  = "Year",
    values_to = "Debt"
  ) %>%
  mutate(
    Year     = as.integer(Year),
    debt_gdp = readr::parse_number(Debt),  # % of GDP
    COUNTRY  = COUNTRY                      # keep the IMF country name
  ) %>%
  select(COUNTRY, Year, debt_gdp) %>%
  filter(!is.na(Year), !is.na(debt_gdp)) %>%
  arrange(Year)

# sanity check
nrow(cad_debt)
range(cad_debt$Year)
head(cad_debt)
distinct(cad_debt, COUNTRY)
```

```{r}
ca_panel <- reduce(
  list(cad_pop, cad_mfp, cad_gdp, cad_debt),
  ~ inner_join(.x, .y, by = c("COUNTRY", "Year"))
) %>%
  filter(COUNTRY == "Canada") %>%
  arrange(Year)
ca <- ca_panel %>%
  arrange(Year) %>%
  mutate(
    gdp_per_capita = (gdp_level * 1e6) / Population,
    log_gdppc      = log(gdp_per_capita),
    log_mfp        = log(MFP_index),
    gdp_growth     = 100 * (gdp_level / dplyr::lag(gdp_level) - 1)
  ) %>%
  filter(!is.na(debt_gdp), !is.na(gdp_growth)) %>%
  arrange(Year)
```

```{r}
canada <- ca %>%
  select(COUNTRY, Year, gdp_growth, debt_gdp,
         gdp_per_capita, log_gdppc, log_mfp) %>%
  arrange(Year) %>%
  mutate(
    gdppc_growth = 100 * (gdp_per_capita / dplyr::lag(gdp_per_capita) - 1)
  )
```

```{r}
canada_master <- ca %>%
  arrange(Year) %>%
  mutate(
    # Growth rates
    delta_mfp   = 100 * (log_mfp - dplyr::lag(log_mfp, 1)),
    debt_lag1   = dplyr::lag(debt_gdp, 1),
    debt_lag2   = dplyr::lag(debt_gdp, 2),
    debt_lag1_2 = debt_lag1^2,
    debt_lag2_2 = debt_lag2^2,

    # Crisis dummies
    d_0809 = ifelse(Year %in% c(2008, 2009), 1, 0),
    d_2021 = ifelse(Year %in% c(2020, 2021), 1, 0)
  )
```

```{r}
canada_delta <- canada %>%
  arrange(Year) %>%
  mutate(
    # MFP growth in percentage points
    delta_mfp = 100 * (log_mfp - dplyr::lag(log_mfp, 1))
  ) %>%
  drop_na(delta_mfp, gdp_growth, debt_gdp, log_mfp)
```

```{r}
canada_dum <- canada_delta %>%
  mutate(
    d_0809 = ifelse(Year %in% c(2008, 2009), 1, 0),
    d_2021 = ifelse(Year %in% c(2020, 2021), 1, 0)
  )
```

```{r}
key_years <- cad_debt %>%
  filter(Year %in% c(1990, 1995, 2000, 2007, 2008, 2009, 2019, 2020, 2021, 2023)) %>%
  select(Year, debt_gdp) %>%
  arrange(Year)

kable(key_years, 
      col.names = c("Year", "Debt-to-GDP (%)"),
      align = c("l", "r"))
```


# Time Series Graphs:

```{r}
ggplot(ca, aes(x = Year)) + 
  geom_line(aes(y = gdp_growth * 10, color = "GDP Growth (%)"), linewidth = 1.2) +
  scale_y_continuous( sec.axis = sec_axis(~ . / 5, name = "GDP Growth (%)")
  ) +
  scale_color_manual(values = c("GDP Growth (%)" = "#D55E00")) +
    labs(
    title = "Canada: GDP Growth",
    x = "Year",
    color = ""
  ) +
  theme_minimal(base_size = 9) +
  theme(legend.position = "bottom")
```


```{r}
ggplot(ca, aes(x = Year)) +
  # Add shaded regions for crisis periods FIRST (so they're in background)
  annotate("rect", xmin = 2008, xmax = 2009, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "gray") +
  annotate("rect", xmin = 2020, xmax = 2021, ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "gray") +
  
  # Your existing lines
  geom_line(aes(y = debt_gdp, color = "Debt-to-GDP (%)"), linewidth = 1.2) +
  geom_line(aes(y = gdp_growth * 10, color = "GDP Growth (%)"), linewidth = 1.2) +
  
  scale_y_continuous(
    name = "Debt-to-GDP (%)",
    sec.axis = sec_axis(~ . / 5, name = "GDP Growth (%)")
  ) +
  scale_color_manual(values = c("Debt-to-GDP (%)" = "#0072B2",
                                "GDP Growth (%)" = "#D55E00")) +
  labs(
    title = "Canada: Debt-to-GDP & GDP Growth",
    x = "Year",
    color = ""
  ) +
  theme_minimal(base_size = 9) +
  theme(legend.position = "bottom")
```
```{r}
ca_pop <- ca %>%
  arrange(Year) %>%
  mutate(
    pop_growth = 100 * (Population / dplyr::lag(Population) - 1)
  ) %>%
  filter(!is.na(pop_growth), !is.na(gdp_growth))

ggplot(ca_pop, aes(x = Year)) +
  geom_line(aes(y = pop_growth, color = "Population growth (%)"), linewidth = 1.2) +
  geom_line(aes(y = gdp_growth, color = "GDP growth (%)"), linewidth = 1.2) +
  scale_y_continuous(
    name = "Population growth (%)",
    sec.axis = sec_axis(~ ., name = "GDP growth (%)")  # same scale on both axes
  ) +
  scale_color_manual(
    values = c(
      "Population growth (%)" = "#0072B2",
      "GDP growth (%)"        = "#D55E00"
    )
  ) +
  labs(
    title = "Canada: Population Growth vs GDP Growth",
    x = "Year",
    color = ""
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")
```

```{r}
ca_mfp <- ca %>%
  arrange(Year) %>%
  mutate(
    mfp_growth = 100 * (log_mfp - dplyr::lag(log_mfp))  # ≈ % change in MFP
  ) %>%
  filter(!is.na(mfp_growth), !is.na(gdp_growth))

ggplot(ca_mfp, aes(x = Year)) +
  geom_line(aes(y = mfp_growth, color = "MFP growth (%)"), linewidth = 1.2) +
  geom_line(aes(y = gdp_growth, color = "GDP growth (%)"), linewidth = 1.2) +
  scale_y_continuous(
    name = "MFP growth (%)",
    sec.axis = sec_axis(~ ., name = "GDP growth (%)")
  ) +
  scale_color_manual(
    values = c(
      "MFP growth (%)" = "#0072B2",
      "GDP growth (%)" = "#D55E00"
    )
  ) +
  labs(
    title = "Canada: MFP Growth vs GDP Growth",
    x = "Year",
    color = ""
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")
```
# ADF Test

```{r}
library(urca)

adf_ur <- function(x, max_lags = 6, type = c("none","drift","trend")) {
  type <- match.arg(type)
  # choose lag length by AIC over 0..max_lags
  fits <- lapply(0:max_lags, \(L) ur.df(x, type = type, lags = L, selectlags = "Fixed"))
  aics <- sapply(fits, \(m) AIC(lm(m@testreg)))
  best <- fits[[which.min(aics)]]
  list(best = best, aic = aics)
}

# debt and log mfp: usually include drift or trend
debt_adf_drift <- adf_ur(ca$debt_gdp, type="drift")
mfp_adf_trend  <- adf_ur(ca$log_mfp,  type="trend")

# growth rates: usually "none" or "drift"
gdpgr_adf_none <- adf_ur(ca$gdp_growth, type="none")
dmfp_adf_none  <- adf_ur(100*diff(ca$log_mfp), type="none")

summary(debt_adf_drift$best)
summary(mfp_adf_trend$best)
summary(gdpgr_adf_none$best)
summary(dmfp_adf_none$best)
```

# Baseline Regression 


```{r}
canada_baseline <- canada %>%
  arrange(Year) %>%
  mutate(
    delta_mfp = 100 * (log_mfp - dplyr::lag(log_mfp, 1))
  ) %>%
  drop_na(gdp_growth, debt_gdp, delta_mfp)

m_baseline_nolag <- lm(
  gdp_growth ~ debt_gdp + delta_mfp,
  data = canada_baseline
)
m_baseline_nolag_quad <- lm(
  gdp_growth ~ debt_gdp + I(debt_gdp^2) + delta_mfp,
  data = canada_baseline
)

stargazer(
  m_baseline_nolag, m_baseline_nolag_quad,
  type = "text",
  title = "Canada: Baseline GDP Growth Regressions (No Lag)",
  column.labels = c("Linear", "Quadratic"),
  dep.var.labels = "Real GDP Growth (%)",
  covariate.labels = c(
    "Debt-to-GDP (t)",
    "Debt-to-GDP² (t)",
    "Δ log(MFP)"
  ),
  omit.stat = c("f", "ser"),
  digits = 3,
  no.space = TRUE
)
```

Turning Points
```{r}
b1 <- coef(m_baseline_nolag_quad)["debt_gdp"]
b2 <- coef(m_baseline_nolag_quad)["I(debt_gdp^2)"]

# turning point
tp_baseline <- -b1 / (2 * b2)
tp_baseline
```


Plotting the baseline:
```{r}
debt_seq <- seq(
  min(canada_baseline$debt_gdp),
  max(canada_baseline$debt_gdp),
  length.out = 100
)

# Hold MFP growth at its sample mean
mfp_mean <- mean(canada_baseline$delta_mfp, na.rm = TRUE)

# Prediction dataframe
pred_df <- data.frame(
  debt_gdp = debt_seq,
  delta_mfp = mfp_mean
)

# Predicted GDP growth from quadratic model
pred_df$gdp_pred <- predict(
  m_baseline_nolag_quad,
  newdata = pred_df
)
ggplot(pred_df, aes(x = debt_gdp, y = gdp_pred)) +
  geom_line(size = 1.1) +
  geom_vline(
    xintercept = tp_baseline,
    linetype = "dashed"
  ) +
  labs(
    title = "Fitted Relationship Between Debt-to-GDP and GDP Growth",
    subtitle = "Quadratic specification with Δlog(MFP) held at its mean",
    x = "Debt-to-GDP Ratio (%)",
    y = "Predicted Real GDP Growth (%)"
  ) +
  annotate(
  "text",
  x = tp_baseline,
  y = max(pred_df$gdp_pred),
  label = paste0("Turning point ≈ ", round(tp_baseline, 1), "%"),
  hjust = -0.1,
  vjust = 2
) + 
  theme_minimal()

```

## Baseline Regression Tests:

Joint F-Test on Debt Terms:

```{r}
linearHypothesis(
  m_baseline_nolag_quad,
  c("debt_gdp = 0", "I(debt_gdp^2) = 0"),
  vcov = vcovHAC(m_baseline_nolag_quad)
)
```
RESET Test:

```{r}
resettest(m_baseline_nolag, power = 2:3, type = "fitted")
resettest(m_baseline_nolag_quad, power = 2:3, type = "fitted")
```
Heteroskedasticity Test:

```{r}
bptest(m_baseline_nolag_quad)
```
Serial Correlation Test:

```{r}
bgtest(m_baseline_nolag_quad, order = 2)
```
HAC Inference Test:

```{r}
coeftest(m_baseline_nolag_quad, vcov = vcovHAC(m_baseline_nolag_quad))
```


## 1 Year Debt Lag

```{r}
canada_delta <- canada %>%
  arrange(Year) %>%
  mutate(
    delta_mfp   = 100 * (log_mfp - dplyr::lag(log_mfp, 1)),
    debt_lag1   = dplyr::lag(debt_gdp, 1),
    debt_lag1_2 = debt_lag1^2
  ) %>%
  drop_na(delta_mfp, gdp_growth, debt_lag1, debt_lag1_2)
```

```{r}
m_dynamic_lag <- lm(
  gdp_growth ~ debt_lag1 + debt_lag1_2 + delta_mfp,
  data = canada_delta
)
stargazer(
  m_dynamic_lag,
  type = "text",
  title = "Canada: GDP Growth with Lagged Debt-to-GDP",
  dep.var.labels = "Real GDP Growth (%)",
  covariate.labels = c(
    "Lagged Debt-to-GDP (t−1)",
    "Lagged Debt-to-GDP Squared (t−1)",
    "Δ log(MFP)"
  ),
  omit.stat = c("f", "ser"),
  digits = 3,
  no.space = TRUE
)
```

Turning Points:
```{r}
b <- coef(m_dynamic_lag)

tp <- -b["debt_lag1"] / (2 * b["debt_lag1_2"])
tp
```


## 2 Year Debt Lag

```{r}
canada_delta <- canada %>%
  arrange(Year) %>%
  mutate(
    delta_mfp   = 100 * (log_mfp - dplyr::lag(log_mfp, 1)),
    debt_lag2   = dplyr::lag(debt_gdp, 2),
    debt_lag2_2 = debt_lag2^2
  ) %>%
  drop_na(delta_mfp, gdp_growth, debt_lag2, debt_lag2_2)
```

```{r}
m_dynamic_lag2 <- lm(
  gdp_growth ~ debt_lag2 + debt_lag2_2 + delta_mfp,
  data = canada_delta
)

stargazer(
  m_dynamic_lag2,
  type = "text",
  title = "Canada: GDP Growth with 2-Year Lagged Debt-to-GDP",
  dep.var.labels = "Real GDP Growth (%)",
  covariate.labels = c(
    "Debt-to-GDP (t−2)",
    "Debt-to-GDP Squared (t−2)",
    "Δ log(MFP)"
  ),
  omit.stat = c("f", "ser"),
  digits = 3,
  no.space = TRUE
)
```

```{r}
b <- coef(m_dynamic_lag2)

tp <- -b["debt_lag2"] / (2 * b["debt_lag2_2"])
tp
```


## Graphing the Lag Effects:

```{r}
canada_lags_plot <- canada_master %>%
  select(Year, gdp_growth, debt_gdp, debt_lag1, debt_lag2) %>%
  drop_na()

ggplot(canada_lags_plot, aes(x = Year)) +
  geom_line(aes(y = gdp_growth, color = "GDP Growth"), linewidth = 1.2) +
  geom_line(aes(y = debt_gdp / 5, color = "Debt-to-GDP (t) / 5"),
            linewidth = 1.2, linetype = "solid") +
  labs(
    title = "Canada: GDP Growth and Current Debt-to-GDP",
    x = "Year",
    y = "Scaled Values",
    color = ""
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

ggplot(canada_lags_plot, aes(x = Year)) +
  geom_line(aes(y = gdp_growth, color = "GDP Growth"), linewidth = 1.2) +
  geom_line(aes(y = debt_lag1 / 5, color = "Debt-to-GDP (t−1) / 5"),
            linewidth = 1.2, linetype = "solid") +
  labs(
    title = "Canada: GDP Growth and 1-Year Lagged Debt-to-GDP",
    x = "Year",
    y = "Scaled Values",
    color = ""
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
ggplot(canada_lags_plot, aes(x = Year)) +
  geom_line(aes(y = gdp_growth, color = "GDP Growth"), linewidth = 1.2) +
  geom_line(aes(y = debt_lag2 / 5, color = "Debt-to-GDP (t−2) / 5"),
            linewidth = 1.2, linetype = "solid") +
  labs(
    title = "Canada: GDP Growth and 2-Year Lagged Debt-to-GDP",
    x = "Year",
    y = "Scaled Values",
    color = ""
  ) +
  theme_linedraw() +
  theme(legend.position = "bottom")

```

## Lagged Interaction Terms:

```{r}
canada_me <- canada_dum %>%
  arrange(Year) %>%
  mutate(
    debt_lag1   = dplyr::lag(debt_gdp, 1),
    debt_lag1_2 = debt_lag1^2,

    # interactions from the board idea
    dL1_x_dmfp  = debt_lag1 * delta_mfp,
    dL1_x_0809  = debt_lag1 * d_0809,
    dL1_x_2021  = debt_lag1 * d_2021
  ) %>%
  drop_na(gdp_growth, debt_lag1, debt_lag1_2, delta_mfp, d_0809, d_2021)

```

```{r}
m_board <- lm(
  gdp_growth ~ debt_lag1 + debt_lag1_2 +
    delta_mfp + d_0809 + d_2021 +
    debt_lag1:delta_mfp + debt_lag1:d_0809 + debt_lag1:d_2021,
  data = canada_me
)

nw_vcov <- sandwich::vcovHAC(m_board)
lmtest::coeftest(m_board, vcov = nw_vcov)

stargazer(
  m_board,
  type = "text",
  title = "Canada: Interactive Model (Lagged Debt, ΔMFP, Crisis Dummies)",
  dep.var.labels = "GDP Growth (%)",
  omit.stat = c("f","ser"),
  digits = 3,
  no.space = TRUE
)
```

### Marginal Effects Computation

```{r}
b <- coef(m_board)

canada_me <- canada_me %>%
  mutate(
    me_debt_lag1 =
      b["debt_lag1"] +
      2 * b["debt_lag1_2"] * debt_lag1 +
      b["debt_lag1:delta_mfp"] * delta_mfp +
      b["debt_lag1:d_0809"] * d_0809 +
      b["debt_lag1:d_2021"] * d_2021
  )
```

```{r}
V <- sandwich::vcovHAC(m_board)

# plug in numeric sample means (computed from the data used in m_board)
xbar_d   <- mean(canada_me$debt_lag1,   na.rm = TRUE)
xbar_dm  <- mean(canada_me$delta_mfp,  na.rm = TRUE)
xbar_0809<- mean(canada_me$d_0809,     na.rm = TRUE)
xbar_2021<- mean(canada_me$d_2021,     na.rm = TRUE)

expr <- paste0(
  "debt_lag1 + 2*debt_lag1_2*", xbar_d,
  " + `debt_lag1:delta_mfp`*", xbar_dm,
  " + `debt_lag1:d_0809`*", xbar_0809,
  " + `debt_lag1:d_2021`*", xbar_2021
)

me_avg <- car::deltaMethod(m_board, g = expr, vcov. = V)

me_est <- as.numeric(me_avg[1])
me_se  <- as.numeric(me_avg[2])

# helper: delta-method ME at chosen (d, dmfp, d0809, d2021)
me_at <- function(d, dmfp, d0809, d2021, label) {
  expr <- paste0(
    "debt_lag1 + 2*debt_lag1_2*", d,
    " + `debt_lag1:delta_mfp`*", dmfp,
    " + `debt_lag1:d_0809`*", d0809,
    " + `debt_lag1:d_2021`*", d2021
  )
  dm <- car::deltaMethod(m_board, g = expr, vcov. = V)
  est <- as.numeric(dm[1]); se <- as.numeric(dm[2])
  t   <- est / se
  p   <- 2 * pt(abs(t), df = df.residual(m_board), lower.tail = FALSE)

  data.frame(
    Scenario = label,
    Debt_eval = d,
    dMFP_eval = dmfp,
    d_0809 = d0809,
    d_2021 = d2021,
    ME = est,
    `HAC SE` = se,
    `t-stat` = t,
    `p-value` = p,
    `95% CI (low)` = est - 1.96 * se,
    `95% CI (high)` = est + 1.96 * se
  )
}

d_bar    <- mean(canada_me$debt_lag1, na.rm = TRUE)
dmfp_bar <- mean(canada_me$delta_mfp, na.rm = TRUE)

me_tbl <- bind_rows(
  me_at(d_bar, dmfp_bar, 0, 0, "Normal year (means; no crisis)"),
  me_at(d_bar, dmfp_bar, 1, 0, "2008–09 crisis year (means)"),
  me_at(d_bar, dmfp_bar, 0, 1, "2020–21 crisis year (means)"),
  me_at(d_bar, 0,        0, 0, "Normal year (ΔMFP = 0; no crisis)")
)
me_ultra_clean <- me_tbl %>%
  dplyr::select(
    Scenario,
    ME,
    X95..CI..low.,
    X95..CI..high.
  )

stargazer(
  me_ultra_clean,
  type = "text",
  title = "Canada: Marginal Effect of Lagged Debt-to-GDP on GDP Growth",
  summary = FALSE,
  digits = 4,
  rownames = FALSE,
  covariate.labels = c(
    "Scenario",
    "Marginal Effect",
    "95% CI (Lower)",
    "95% CI (Upper)"
  ),
  notes = "Marginal effects computed using the delta method with HAC standard errors."
)
```

Plotting:

```{r}
ggplot(canada_me, aes(x = Year, y = me_debt_lag1)) +
  geom_line(linewidth = 1.1) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Canada: Marginal Effect of Lagged Debt-to-GDP on GDP Growth",
    x = "Year",
    y = "∂ŷ/∂Debt(t−1)"
  ) +
  theme_minimal(base_size = 13)

```

#GDPPC Regression

## Baseline

```{r}
canada_gdppc <- canada_dum %>%
  drop_na(gdppc_growth, delta_mfp)

m_gdppc <- lm(
  gdppc_growth ~ debt_gdp + I(debt_gdp^2) + delta_mfp + d_0809 + d_2021,
  data = canada_gdppc
)

stargazer(
  m_gdppc,
  type = "text",
  title = "Canada: GDP per Capita Growth Regression with Δlog(MFP) and Recession Dummies",
  dep.var.labels = "GDP per Capita Growth (%)",
  column.labels = "Base Model",
  covariate.labels = c(
    "Debt-to-GDP",
    "Debt-to-GDP²",
    "Δlog(MFP)",
    "Dummy: 2008–2009",
    "Dummy: 2020–2021"
  ),
  omit.stat = c("f", "ser"),
  digits = 3,
  no.space = TRUE
)
```

Turning Points:

```{r}
b1 <- coef(m_gdppc)["debt_gdp"]
b2 <- coef(m_gdppc)["I(debt_gdp^2)"]

turning_point <- -b1 / (2 * b2)
turning_point
```

#One Year Lagged 

```{r}
canada_gdppc_lag <- canada_dum %>%
  arrange(Year) %>%
  mutate(
    gdppc_growth_lag1 = dplyr::lag(gdppc_growth, 1),
    debt_lag1         = dplyr::lag(debt_gdp, 1),
    debt_lag1_2       = debt_lag1^2
  ) %>%
  drop_na(
    gdppc_growth, gdppc_growth_lag1,
    debt_lag1, debt_lag1_2,
    delta_mfp, d_0809, d_2021
  )

m_gdppc_lag1 <- lm(
  gdppc_growth ~
    debt_lag1 + debt_lag1_2 +
    delta_mfp + d_0809 + d_2021,
  data = canada_gdppc_lag
)
stargazer(
  m_gdppc_lag1,
  type = "text",
  title = "Canada: GDP per Capita Growth with Lagged Debt-to-GDP",
  dep.var.labels = "GDP per Capita Growth (%)",
  covariate.labels = c(
    "Debt-to-GDP (t−1)",
    "Debt-to-GDP² (t−1)",
    "Δ log(MFP)",
    "Dummy: 2008–2009",
    "Dummy: 2020–2021"
  ),
  omit.stat = c("f", "ser"),
  digits = 3,
  no.space = TRUE
)

```

# Reasoning for Change in Log(MFP) in Baseline Regression:


```{r}
canada <- ca %>%
  select(COUNTRY, Year, gdp_growth, debt_gdp,
         gdp_per_capita, log_gdppc, log_mfp) %>%
  arrange(Year) %>%
  mutate(
    gdppc_growth = 100 * (gdp_per_capita / dplyr::lag(gdp_per_capita) - 1)
  )
```

```{r}
canada_controls <- canada %>%
  drop_na(gdp_growth, debt_gdp, log_mfp)

m1 <- lm(
  gdp_growth ~ debt_gdp + log_mfp,
  data = canada_controls
)

m2 <- lm(
  gdp_growth ~ debt_gdp + I(debt_gdp^2) + log_mfp,
  data = canada_controls
)

tp <- -coef(m2)["debt_gdp"] / (2 * coef(m2)["I(debt_gdp^2)"])

stargazer(
  m1, m2,
  type = "text",
  title = "Canada: Growth Regressions with MFP",
  dep.var.labels = "Real GDP Growth (%)",
  column.labels  = c("Baseline", "Quadratic Debt"),
  covariate.labels = c(
    "Debt-to-GDP",
    "Debt-to-GDP²",
    "log(MFP)"
  ),
  omit.stat = c("f", "ser"),
  digits = 3,
  no.space = TRUE
)
```

Plotting the Results:

```{r}
pred_canada <- canada_controls %>%
  mutate(
    pred_m1 = predict(m1, newdata = canada_controls),
    pred_m2 = predict(m2, newdata = canada_controls)
  )

ggplot(canada_controls, aes(x = debt_gdp, y = gdp_growth)) +
  geom_point(alpha = 0.6) +
  geom_line(
    data = pred_canada,
    aes(y = pred_m1),
    linetype = "dashed"
  ) +
  geom_line(
    data = pred_canada,
    aes(y = pred_m2)
  ) +
  geom_vline(
    xintercept = as.numeric(tp),
    linetype = "dotted"
  ) +
  annotate(
    "text",
    x = as.numeric(tp),
    y = min(canada_controls$gdp_growth, na.rm = TRUE),
    label = paste0("TP ≈ ", round(tp, 1), "%"),
    vjust = -0.5,
    angle = 90
  ) +
  labs(
    x = "Debt-to-GDP ratio (%)",
    y = "Real GDP growth (%)",
    title = "Canada: GDP Growth vs Debt-to-GDP",
    subtitle = "Points = data; dashed = linear fit; solid = quadratic fit; dotted = turning point"
  ) +
  theme_minimal()
```

## Lagged MFP versus Non Lagged MFP

```{r}
canada_lag <- canada %>%
  arrange(Year) %>%
  mutate(
    mfp_lag  = dplyr::lag(log_mfp, 1),
    debt_lag = dplyr::lag(debt_gdp, 1)
  ) %>%
  drop_na(gdp_growth, debt_gdp, mfp_lag)

m_lag1 <- lm(
  gdp_growth ~ debt_gdp + I(debt_gdp^2) + mfp_lag,
  data = canada_lag
)

stargazer(
  m_lag1,
  type = "text",
  title = "Canada: Lagged MFP Regression (Dependent Variable: GDP Growth)",
  dep.var.labels = "GDP Growth (%)",
  column.labels = "Lagged MFP Model",
  covariate.labels = c(
    "Debt-to-GDP",
    "Debt-to-GDP²",
    "Lagged log(MFP)"
  ),
  omit.stat = c("f", "ser"),
  digits = 3,
  no.space = TRUE
)
```

## Change in MFP:

```{r}
canada_delta <- canada %>%
  arrange(Year) %>%
  mutate(
    # MFP growth in percentage points
    delta_mfp = 100 * (log_mfp - dplyr::lag(log_mfp, 1))
  ) %>%
  drop_na(delta_mfp, gdp_growth, debt_gdp, log_mfp)

m_dynamic <- lm(
  gdp_growth ~ debt_gdp + I(debt_gdp^2) + delta_mfp,
  data = canada_delta
)

stargazer(
  m2, m_dynamic,
  type = "text",
  title = "Canada: Comparing Productivity Specifications",
  column.labels = c("Quadratic Debt + log(MFP)", "Quadratic Debt + Δlog(MFP)"),
  dep.var.labels = "Real GDP Growth (%)",
  covariate.labels = c(
    "Debt-to-GDP",
    "Debt-to-GDP²",
    "log(MFP)",
    "Δlog(MFP)"
  ),
  omit.stat = c("f", "ser"),
  digits = 3,
  no.space = TRUE
)
```
#Rolling Analysis

## Rolling 12 Year Turning Point Analysis:

```{r}
roll_tp <- function(df, window_size = 12) {

  years <- df$Year

  results <- lapply(1:(length(years) - window_size + 1), function(i) {

    sub <- df %>%
      filter(Year >= years[i], Year < years[i] + window_size)

    if (nrow(sub) < window_size - 2) return(NULL)

    fit <- tryCatch(
      lm(gdp_growth ~ debt_gdp + I(debt_gdp^2), data = sub),
      error = function(e) NULL
    )
    if (is.null(fit)) return(NULL)

    co <- coef(fit)
    b1 <- co["debt_gdp"]
    b2 <- co["I(debt_gdp^2)"]
    tp <- -b1 / (2 * b2)

    tp_se <- {
      vc <- vcov(fit)
      sqrt(
        vc["debt_gdp","debt_gdp"] / (4 * b2^2) +
        (b1^2 * vc["I(debt_gdp^2)","I(debt_gdp^2)"]) / (4 * b2^4) -
        (b1 * vc["debt_gdp","I(debt_gdp^2)"]) / (2 * b2^3)
      )
    }

    tibble(
      start = years[i],
      end   = years[i] + window_size - 1,
      tp     = tp,
      tp_low = tp - 1.96 * tp_se,
      tp_high = tp + 1.96 * tp_se,
      beta2 = b2
    )
  })

  bind_rows(results)
}

tp_ca <- roll_tp(ca, window_size = 12)
```

Plotting the results:

```{r}
ggplot(tp_ca, aes(x = end, y = tp)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Canada: Rolling Turning Point (12-Year Windows)",
    x = "End Year",
    y = "Turning Point (% Debt-to-GDP)"
  ) +
  theme_minimal(base_size = 13)
```

## Split - Sample Analysis (\>2007 vs 2008\<)

```{r}
ca_pre  <- ca %>% filter(Year <= 2007)
ca_post <- ca %>% filter(Year >= 2008)

m_pre  <- lm(gdp_growth ~ debt_gdp + I(debt_gdp^2), data = ca_pre)
m_post <- lm(gdp_growth ~ debt_gdp + I(debt_gdp^2), data = ca_post)
```

```{r}
tp_pre  <- -coef(m_pre)["debt_gdp"] / (2 * coef(m_pre)["I(debt_gdp^2)"])
tp_post <- -coef(m_post)["debt_gdp"] / (2 * coef(m_post)["I(debt_gdp^2)"])
```

Plotting:

```{r}
ggplot() +
  geom_point(data = ca_pre, aes(debt_gdp, gdp_growth), color = "#0072B2") +
  geom_line(data = ca_pre, aes(debt_gdp, predict(m_pre)), color = "#0072B2") +
  geom_point(data = ca_post, aes(debt_gdp, gdp_growth), color = "#D55E00") +
  geom_line(data = ca_post, aes(debt_gdp, predict(m_post)), color = "#D55E00") +
  theme_minimal(base_size = 13) +
  labs(
    title = "Debt–Growth Curve Before vs After 2008",
    x = "Debt-to-GDP (%)",
    y = "GDP Growth (%)"
  )
```

## Rolling 10 Year Correlation:

```{r}
ca_corr <- ca %>%
  arrange(Year) %>%
  mutate(
    roll_corr = rollapply(
      data = cbind(gdp_growth, debt_gdp),
      width = 10,
      FUN = function(z) cor(z[,1], z[,2], use = "complete.obs"),
      by.column = FALSE,
      fill = NA,
      align = "right"
    )
  )
```

```{r}
ggplot(ca_corr, aes(Year, roll_corr)) +
  geom_line(color = "#0072B2", linewidth = 1.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Canada: Rolling 10-Year Correlation Between Debt and Growth",
    x = "Year",
    y = "Correlation"
  ) +
  theme_minimal(base_size = 1)
```


# MFP "shock" Series:

Constructing the dataset:

```{r}
canada_cf <- canada_master %>%
  arrange(Year) %>%
  mutate(
    debt_lag1 = dplyr::lag(debt_gdp, 1),
    debt_lag2 = dplyr::lag(debt_gdp, 2)
  ) %>%
  drop_na(delta_mfp, gdp_growth, debt_lag1)
```

The baseline linear regression with MFP on the LS:

```{r}
mfp_fs <- lm(delta_mfp ~ debt_lag1, data = canada_cf)

canada_cf <- canada_cf %>%
  mutate(
    mfp_hat   = fitted(mfp_fs),
    mfp_shock = resid(mfp_fs)  
  )
stargazer(
  mfp_fs,
  type = "text",
  title = "First-Stage Regression: Determinants of MFP Growth",
  dep.var.labels = "Δlog(MFP) (%)",
  covariate.labels = c(
    "Lagged Debt-to-GDP (t−1)"
  ),
  omit.stat = c("f","ser"),
  digits = 3,
  no.space = TRUE,
  notes = "Residuals from this regression are used as productivity shocks in the GDP growth regressions."
)
```


## MFP driven by lagged debt only and dummies:

```{r}
mfp_fs_rich <- lm(delta_mfp ~ debt_lag1 + d_0809 + d_2021, data = canada_cf)

canada_cf <- canada_cf %>%
  mutate(
    mfp_hat   = fitted(mfp_fs_rich),
    mfp_shock = resid(mfp_fs_rich)
  )

stargazer(
  mfp_fs_rich,
  type = "text",
  title = "First-Stage Regression (Rich): Determinants of MFP Growth",
  dep.var.labels = "Δlog(MFP) (%)",
  covariate.labels = c(
    "Lagged Debt-to-GDP (t−1)",
    "Dummy: 2008–2009",
    "Dummy: 2020–2021"
  ),
  omit.stat = c("f","ser"),
  digits = 3,
  no.space = TRUE,
  notes = "Residuals from this regression (mfp_shock) are used as productivity shocks in the GDP growth regressions."
)
```


Now, showing the growth on debt + MFP shocks 

```{r}

m_cf_lin <- lm(gdp_growth ~ debt_lag1 + mfp_shock, data = canada_cf)
m_cf_quad <- lm(gdp_growth ~ debt_lag1 + I(debt_lag1^2) + mfp_shock, data = canada_cf)

stargazer(
  m_cf_lin, m_cf_quad,
  type = "text",
  title = "Canada: Control-Function Style Regressions (MFP Shock Residual)",
  column.labels = c("Linear debt", "Quadratic debt"),
  dep.var.labels = "Real GDP Growth (%)",
  covariate.labels = c(
    "Debt-to-GDP (t−1)",
    "Debt-to-GDP² (t−1)",
    "MFP shock (first-stage residual)"
  ),
  omit.stat = c("f","ser"),
  digits = 3,
  no.space = TRUE
)
```

Turning Points:

```{r}
b1 <- coef(m_cf_quad)["debt_lag1"]
b2 <- coef(m_cf_quad)["I(debt_lag1^2)"]
tp_cf <- -b1 / (2*b2)
tp_cf
```


## Adding the shock MFP into the interactive model:

```{r}
m_board_cf <- lm(
  gdp_growth ~ debt_lag1 + I(debt_lag1^2) +
    mfp_shock + d_0809 + d_2021 +
    debt_lag1:mfp_shock + debt_lag1:d_0809 + debt_lag1:d_2021,
  data = canada_cf
)

V_hac  <- vcovHAC(m_board_cf)
se_hac <- sqrt(diag(V_hac))


stargazer(
  m_board_cf,
  type = "text",
  title = "Canada: Interactive Model with MFP Shock (HAC Standard Errors)",
  dep.var.labels = "Real GDP Growth (%)",
  covariate.labels = c(
    "Debt-to-GDP (t−1)",
    "Debt-to-GDP² (t−1)",
    "MFP shock (residual)",
    "Dummy: 2008–2009",
    "Dummy: 2020–2021",
    "Debt(t−1) × MFP shock",
    "Debt(t−1) × 2008–2009",
    "Debt(t−1) × 2020–2021"
  ),
  se = list(se_hac),
  omit.stat = c("f", "ser"),
  digits = 3,
  no.space = TRUE,
  notes = "HAC standard errors from vcovHAC()."
)
```

Graphing:

```{r}
ggplot(canada_cf, aes(x = Year)) +
  geom_line(aes(y = delta_mfp, color = "Actual ΔMFP"), linewidth = 1.1) +
  geom_line(aes(y = mfp_hat, color = "Predicted ΔMFP"), linewidth = 1.1, linetype = "dashed") +
  labs(
    title = "Canada: Actual vs Predicted MFP Growth",
    subtitle = "Predicted values from first-stage regression",
    x = "Year",
    y = "Δlog(MFP) (%)",
    color = ""
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")
```


Time Series of MFP Shocks:

```{r}
ggplot(canada_cf, aes(x = Year, y = mfp_shock)) +
  geom_line(linewidth = 1.1) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Canada: Productivity Shocks (First-Stage Residuals)",
    x = "Year",
    y = "MFP shock (Δlog(MFP) residual)"
  ) +
  theme_minimal(base_size = 13)
```
